---
title: "ColorAnalysis"
author: "Shannon E. Ellis"
date: "11/23/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Here we will try to:

1. determine what color palletes people like best  
2. determine which they find most suitable for which graphs  
3. settle the color debate between Jeff and myself  

```{r read in data}
## these data have had duplicates removed
## and emails have been removed
df <- read.csv("Colors.csv", na.strings=c("","NA"))
```

```{r quick-look-at-the-data}
## sex breakdown 
table(droplevels(df$sex))
```

## Palette Preference
Now we'll look at overall preferences by graph type (not broken down by sex).

```{r plotting-preferences2, echo=FALSE}
##plot differentiation of blues
toplot= droplevels(df[,c("GraphingFavorite","GraphingBoxplot","GraphingNetwork")])
  
counts2=c()
for(i in 1:ncol(toplot)){
  counts = table(toplot[,i])/nrow(toplot) 
  counts2 = cbind(counts2,counts)
}
colnames(counts2) <- colnames(toplot)

bright= c(yellow=rgb(255,222,13, maxColorValue=255), 
                    orange=rgb(232,121,12, maxColorValue=255),   
                    red=rgb(255,0,0, maxColorValue=255), 
                    green=rgb(12,189,24, maxColorValue=255), 
                    purple=rgb(148,12,232, maxColorValue=255), 
                    blue=rgb(58,158,234, maxColorValue=255), 
                    pink=rgb(255,78,105, maxColorValue=255), 
                    teal=rgb(59,196,199, maxColorValue=255))
  barplot(t(counts2), beside=T, 
          col=c(bright["green"],bright["purple"],bright["orange"]),ylab="Proportion of Responses",ylim=c(0,0.4), main="Color Preferences by Graph", las=2) 
  legend("topleft", c(colnames(counts2)), pch=15, 
         col=c(bright["green"],bright["purple"],bright["orange"]), 
         bty="n")
```

And, what about an individual's preference from one plot type to another, does that change? Or is their color choice consisitent

```{r preferences-across-graphs}
## do people make the same choice for graphing a network as they do graphing a boxplot?
table(df[,"GraphingFavorite"]==df[,"GraphingBoxplot"])
table(df[,"GraphingFavorite"]==df[,"GraphingNetwork"])
table(df[,"GraphingBoxplot"]==df[,"GraphingNetwork"])
```

```{r preferences-aesthetics}
## are people's aesthetic choices the same as their graphing choices?
table(df[,"GraphingFavorite"]==df[,"AestheticFavorite"])
table(df[,"GraphingBoxplot"]==df[,"AestheticBoxplot"])
table(df[,"GraphingNetwork"]==df[,"AestheticNetwork"])
```

## Color Differentiation

The impetus for this analysis as I recall it:  
Shannon: Well, and this is not a sexist comment, males don't differentiate colors as well as females.  
Jeff: What? Yes that is sexist!  
Shannon: No, it's not. It's a thing.  
Jeff: Are you sure?  
Shannon: I think so...  
Jeff: Let's test it.  

From the quiz, the dots that were referenced in this question (although the sizes and positions of dots were presented differently):

```{r blues, echo=FALSE,fig.width = 9, fig.height = 2}
  blues = c(extralightblue=rgb(237,240,245, maxColorValue=255), 
                    lightblue =rgb(173,185,211, maxColorValue=255),   
                    blue=rgb(99,123,173, maxColorValue=255),   
                    # darkblue=rgb(58,87,149, maxColorValue=255), 
                    darkerblue=rgb(29,52,105, maxColorValue=255),
                    darkestblue=rgb(11,18,36, maxColorValue=255))

par(mfrow=c(1,3), mar=c(1,1,2,1))
flavors=blues
values = c(5,7,9)

for(value in 1:length(values)){
        plot(c(-10, 10), c(-10, 10), xaxt='n', yaxt='n', type='n', xlab='', ylab='', main="")
        xvals = seq(-8, 8, length=values[value]) 
        yvals = runif(values[value], min=-7, max=7)
        points(xvals, yvals, cex=7, pch=21, col='gray70', bg=flavors)
}

```

Note: for the graphs below, the actual number of distinct colors are 5, 7, and 7, respectively.

```{r color-differentiation, echo=FALSE,fig.width = 9}
##plot differentiation of blues
## only include males and females due to sample sizes
## would love to include nonbinary at some point
## just don't have the numbers at this point
sexes = c("Female","Male")
df= df[df$sex %in% sexes,]
df$sex <- droplevels(df$sex)

par(mfrow=c(1,3))
for(i in c("Blues1","Blues2","Blues3")){
  toplot= df[,i]
  counts <- as.matrix(table(toplot,by=df$sex))
  nF<-length(df$sex[df$sex=="Female"])
  nM<-length(df$sex[df$sex=="Male"])
                        
  counts = cbind(counts[,"Female"]/nF,counts[,"Male"]/nM)

  barplot(t(counts), beside=T, 
          col=c(bright["purple"],bright["green"]),ylab="Proportion of Responses",ylim=c(0,1),xlab=paste0("N=",length(df$sex))) 
  legend("topleft", c(paste0("female (N=",nF,")"),paste0("male (N=",nM,")")), pch=15, 
         col=c(bright["purple"],bright["green"]), 
         bty="n")
}
```

## Sex Preferences
Quiz Question: Select any palette in which all the colors included in that palette do not appear distinct to your eyes at first glance. 

```{r breakdown-by-sex}
## breakdown the data by sex
df_M <- df %>% filter(sex=="Male")
df_F <- df %>% filter(sex=="Female")

## females 
length(grep("distinct", df_F$NondistinctPalette))/nrow(df_F)
## males
length(grep("distinct", df_M$NondistinctPalette))/nrow(df_M)

## how about some significance
sexdata <-
matrix(c(length(grep("distinct", df_M$NondistinctPalette)), length(grep("distinct", df_F$NondistinctPalette)), nrow(df_M)-length(grep("distinct", df_M$NondistinctPalette)), nrow(df_F)-length(grep("distinct", df_F$NondistinctPalette))),
       nrow = 2,
       dimnames =
       list(c("Males", "Females"),
            c("Distinct", "Not Distinct")))

fisher.test(sexdata,alternative="less")
```
The percentage who found all of the palletes reasonably distinct: 
 
Females: `r round(length(grep("distinct", df_F$NondistinctPalette))/nrow(df_F)*100,2)`%.

Males: `r round(length(grep("distinct", df_M$NondistinctPalette))/nrow(df_M)*100,2)`%

## Plotting Preferences

For the three following plots, we asked three questions. Proportion of respones for each are plotted:

1. Check the (1) palette you think would be best for graphing.

2. Check the (1) palette you find most appropriate for graphing the boxplots below.

3. Check the (1) palette you find most appropriate for graphing the networks below.
```{r plotting-preferences, echo=FALSE}
##plot differentiation of blues
for(i in c("GraphingFavorite","GraphingBoxplot","GraphingNetwork")){
  toplot= droplevels(df[,i])
  counts <- as.matrix(table(toplot,by=df$sex))
  nF<-length(df$sex[df$sex=="Female"])
  nM<-length(df$sex[df$sex=="Male"])
  N <- nrow(df) 
  
  counts2 = cbind(table(toplot)/length(toplot), counts[,"Female"]/nF,counts[,"Male"]/nM)
  
  barplot(t(counts2), beside=T, 
          col=c("black",bright["purple"],bright["green"]),ylab="Proportion of Responses",ylim=c(0,0.4), main=colnames(df)[i], las=2) 
  legend("topleft", c(paste0("combined (N=",N,")"),paste0("female (N=",nF,")"),paste0("male (N=",nM,")")), pch=15, 
         col=c("black",bright["purple"],bright["green"]), 
         bty="n")
}
```


 